---
title: "Experimental Design Final Project"
author: "David S, Mason"
date: "4/17/2021"
output: html_document
---
This is a script for analyzing bird observations from 42 camera traps monitoring artificial perches above seed traps at 7 paired parcels on Tall Timbers Research Station. These pairs include a parcel burned in Spring of 2020 and another burned the year prior (one-year rough). Each parcel has three traps and the raw data included here runs through April and May.

The overarching hypothesis of my dissertation is that animal resource tracking associated with resources and disturbance
results in plants being disproportionately dispersed toward favorable habitats. In the context of prescribed burns, I predict that the magnet effect of fire will result in increased frugivore observations, which will result in increased seed rain and plant establishment. In this script, I am modeling the first part of that process as my final project for Dr. Phil Hahn's Experimental Design course at University of Florida.

## Set-up the workspace
```{r set-up, warning=FALSE,message=FALSE}
library(tidyverse)
library(lubridate)

brd <- read.csv("PRELIM CAM.csv")
head(brd)
summary(brd)
```

## Data Munging

```{r munging}
# Convert the date into a format the tidyverse trucks with
brd$Date <- mdy(brd$Date)

# Round the date by week to improve figure EDA interpretation
brd$Week <- round_date(brd$Date, unit = 'week')

# Create vector of animals observed which are not birds
not.brds <- c("Bat", "Southern flying squirrel", "Unknown animal",
							"Unknown bat", "Unknown mammal", "Unknown mouse",
							"White-tailed deer", "Southern flying squirrel ")

# Filter out non-birds from the data set
brd.filt <- brd %>% filter(!Species %in% not.brds)

# Drop unknowns we can't assign a functional group to
brd.filt <- brd.filt %>% filter(!Species == "Unknown bird")

# Drop vulture because there is only one scavenger observation
brd.filt <- brd.filt %>% filter(!Species == "Unknown vulture")

# Check to make sure we only have useful values left
spec.list <- as.data.frame(unique(brd.filt$Species))

# Need to make Towhee into Eastern towhee for consistency
levels(brd.filt$Species)[levels(brd.filt$Species) == "Towhee"] <- "Eastern towhee"

# Need to make Eastern Kingbird into Eastern kingbird for consistency
levels(brd.filt$Species)[levels(brd.filt$Species) == "Eastern Kingbird"] <- "Eastern kingbird"
```

## Functional groups

I need to convert the species level data here in two broad functional groups: frugivores and others (insectivores, granivores, scavengers, and predators).


```{r functional groups}
# Make vectors for each functional group
Frugivores <- c("American crow", "Blue jay", "Eastern towhee",
							 "Northern cardinal", "Pileated woodpecker",
							 "Unknown woodpecker", "Eastern bluebird", "Eastern kingbird", 
								"Eastern phoebe",	"Northern mockingbird", "Red-eyed vireo", 
								"Yellow-breasted chat", "Yellow-throated vireo")
Others <- c("Mourning dove", "Blue grosbeak", "Brown-headed cowbird", 
						"Field sparrow", "Carolina wren", "Chuck-will's-widow", "Common yellowthroat",
						"Great crested flycatcher", "Loggerhead shrike")

# Create and fill a column for functional groups
brd.filt.func <- brd.filt %>% 
	mutate(Functional = ifelse(Species %in% Frugivores, "Frugivore",
											ifelse(Species %in% Others, "Other","None")))

# Check to make sure this worked
unique(brd.filt.func$Functional)

# We have one trap that was started earlier than others &
# there may be a true zero at control. Either way it looks messy.

brd.filt.func <- brd.filt.func %>% 
	filter(!Week=='2020-04-12')
```

## Exploratory data analysis

```{r EDA}
# Check out the new data set
summary(brd.filt.func)

# Looking at bird observations by treatment
brd.filt.func %>% group_by(Treatment,Site) %>% summarize(Count = n()) %>% 
	ggplot(aes(x = Treatment, y = Count)) + geom_boxplot()

# Looking by site (looks like an important RV)
brd.filt.func %>% group_by(Site, Treatment) %>% summarize(Count = n()) %>% 
	ggplot(aes(x = Treatment, y = Count)) + geom_col() + facet_wrap(~Site)

# Looking by functional group (looks like a possible covariate)
brd.filt.func %>% group_by(Treatment, Functional) %>% 
	summarize(Count = n()) %>% 
	ggplot(aes(x = Treatment, y = Count, color = Functional)) + 
	geom_col(position = "Dodge") 

# Lets explore the magnet effect with Date/Week (WOAH!)
brd.filt.func %>% 
	group_by(Treatment,Week,Functional) %>% 
	summarize(Count = n()) %>% 
	ggplot(aes(x = Week, y = Count, color = Functional))+
	geom_point()+
	geom_line()+
	facet_wrap(~Treatment)
```

## Summarizing the data

Now that I have a cleaned up version of the data set with the functional role of each observation identified, I need to add up the number of observations for each unique site/trap/date/functional combination.

```{r summarize the data}
# Data to compare all bird observations by treatment
tot.obs <- brd.filt.func %>% 
	group_by(Treatment,Site,Trap,Functional) %>% 
	summarize(Count = n())

# Data to visualize trend related to time
time.obs <- brd.filt.func %>% 
	group_by(Treatment,Site, Functional, Week) %>% 
	summarize(Count = n())
```

## Creating the models

I want to create models to compare the total bird observations for each treatment.

```{r models, warning=FALSE,message=FALSE}
library(lme4)
library(lmerTest)
library(emmeans)
library(car)
library(MuMIn)
library(DHARMa)
library(RVAideMemoire)

# GLM
smpl.mod <- glm(Count~Treatment*Functional,data=tot.obs, family = 'poisson')
summary(smpl.mod)

sim.smpl.mod <- simulateResiduals(fittedModel = smpl.mod, n = 250)
plot(sim.smpl.mod) # warnings

### Test for overdispersion (ie. more variance in the data than expected)
testDispersion(sim.smpl.mod) # overdispersion

### Test for zero-inflation (ie. more zeros in the data than expected)
testZeroInflation(sim.smpl.mod) # not zero-inflated

##### GLMM
nst.mod <- glmer(Count~Treatment*Functional+(1|Site),data=tot.obs, family = 'poisson')
summary(nst.mod)

###### Diagnose GLMM model fit 

sim.nst.mod <- simulateResiduals(fittedModel = nst.mod, n = 250)
plot(sim.nst.mod) # little better than before

overdisp.glmer(nst.mod) # ratio should be 1 (its high)

####### GLM OD
tot.obs$obs <- 1:length(tot.obs$Count) # adding dummy variable as RV

nst.OD.mod <- glmer(Count~Treatment*Functional+(1|Site)+(1|obs),data=tot.obs, family = 'poisson')
summary(nst.OD.mod)

######## Diagnose the OD GLMM
sim.nst.OD.mod <- simulateResiduals(fittedModel = nst.OD.mod, n = 250)
plot(sim.nst.OD.mod) # still doesn't look good

testUniformity(sim.nst.mod) # Failed
testDispersion(sim.nst.mod) # Failed
testZeroInflation(sim.nst.mod) # Failed

######### Negative binomial
nst.nb.mod <- glmer.nb(Count~Treatment*Functional+(1|Site), data=tot.obs)
summary(nst.nb.mod)

########## Diagnose the nb
sim.nst.nb.mod <- simulateResiduals(fittedModel = nst.nb.mod, n = 250)
plot(sim.nst.nb.mod)

########### Poisson with Zero Inflation (note no overdispersion). Is it enough to account for zero inflation without correcting for overdispersion? 
library(glmmTMB) ## new package. similar to lme4 but more flexibility and options

nst.ZI.mod <- glmmTMB(Count~Treatment*Functional+(1|Site), 
                      family=poisson, 
                      zi=~Treatment*Functional, 
                      data=tot.obs)

############ Diagnose the OD GLMM
sim.nst.ZI.mod <- simulateResiduals(fittedModel = nst.ZI.mod, n = 250)
plot(sim.nst.ZI.mod) # still doesn't look good

############# Poisson corrected for zero-inflation AND overdispersion
nst.ZI.OD.mod <- glmmTMB(Count~Treatment*Functional+(1|Site)+(1|obs), 
                     family=poisson, 
                     zi=~Treatment*Functional, 
                     data=tot.obs)
summary(nst.ZI.OD.mod)

############## Diagnose the ZI and OD GLMM 
sim.nst.ZI.OD.mod <- simulateResiduals(fittedModel = nst.ZI.OD.mod, n = 250)
plot(sim.nst.ZI.OD.mod) # Looks good

##############  ZI negative binomial
nst.ZI.nb.mod <- glmmTMB(Count~Treatment*Functional+(1|Site),
                     family=nbinom2, 
                     zi=~Treatment*Functional, 
                     data=tot.obs)

############### Diagnose the model
sim.nst.ZI.nb.mod <- simulateResiduals(fittedModel = nst.ZI.nb.mod, n = 250)
plot(sim.nst.ZI.nb.mod) # Looks decent

```

## Comparing models

```{r comparing models, warning=FALSE,message=FALSE}
summary(smpl.mod) ## simple poisson model
Anova(smpl.mod)  

summary(nst.mod) ## nested glmm
Anova(nst.mod)  

summary(nst.OD.mod)  ## overdispered (OD) nested glmm
Anova(nst.OD.mod)  

summary(nst.nb.mod)  ## nested negative binomial
Anova(nst.nb.mod)

summary(nst.ZI.mod)  ##  zero-inflated (ZI) poisson
Anova(nst.ZI.mod)

summary(nst.ZI.OD.mod)  ## Poisson corrected for ZI AND OD
Anova(nst.ZI.OD.mod)

summary(nst.ZI.nb.mod)  ##  ZI negative binomial
Anova(nst.ZI.nb.mod)

## which model is best?
AIC(smpl.mod,nst.mod,nst.OD.mod,nst.nb.mod,nst.ZI.mod,
		nst.ZI.OD.mod,nst.ZI.nb.mod)

### Poisson corrected for ZI AND OD is the best model
library(glmmTMB)
library(insight)
library(performance)

get_variance(nst.ZI.OD.mod)
r2(nst.ZI.OD.mod)

#### Examine SD associated with random effects 
options (scipen = 999)
VarCorr(nst.ZI.OD.mod) %>% as.data.frame()
SD <- matrix(c("Site","obs","total",0.0000012401,0.8552376549, sum(0.0000012401+0.8552376549)),nrow=3,ncol=2)
colnames(SD) <- c("Groups", "SD")
SD <- as.data.frame(SD)

totSD2 <- VarCorr(fec_odp) %>% as.data.frame() %>% summarize(totSD=sum(vcov)) %>% mutate(totSD=sqrt(totSD))


##### Estimate means with confidence intervals
emmeans(nst.ZI.OD.mod, ~Treatment, type = 'response', bias.adj = T,
				sigma = sqrt(0.8552389))
emmeans(nst.ZI.OD.mod, ~Functional, type = 'response', bias.adj = T,
				sigma = sqrt(0.8552389))

means <- as.data.frame(emmeans(nst.ZI.OD.mod, ~Treatment*Functional, type = 'response', bias.adj = T,
				sigma = sqrt(0.8552389)))

```

## Making a figure representing the means
```{r mean figure, echo=FALSE,warning=FALSE,message=FALSE}
library("ggpubr")
levels(means$Treatment)[levels(means$Treatment)=="B"] <- "Recent burn"
levels(means$Treatment)[levels(means$Treatment)=="C"] <- "One-year rough"


p1 <- ggplot(means, aes(x = Functional, y = rate))+
	geom_linerange(aes(ymin=lower.CL,ymax=upper.CL, col=Functional), 
								 size=1.5)+
	geom_point(aes(col = Functional),size = 8)+
	geom_point(shape = 1,size = 8,colour = "black")+
	theme_bw()+
	theme(text = element_text(size = 22),
				legend.title = element_blank(),
				legend.position = "none",
				plot.title = element_text(hjust = 0.5),
				strip.background = element_blank(),
   			strip.text.y = element_blank(),
				axis.title.x = element_text(face='bold', vjust=-2.5),
				axis.title.y = element_text(face='bold', vjust=3),
				strip.text.x = element_text(size = 18,face ='bold'),
				legend.spacing.x = unit(0.5, 'cm'),
				plot.margin = unit(c(1,1.2,0.9,1.2),"cm"),
				legend.box.spacing = unit(1.2,'cm'),
				axis.ticks.x = element_blank(),
				axis.ticks.y = element_line(size=1.2))+
	scale_color_manual(values = c("#00AFBB", "#E7B800"))+
	scale_y_continuous(breaks=c(0,2,4,6,8,10), limits = c(0,11))+
	ylab("Mean total observations")+
	xlab("Feeding guild")+
	facet_wrap(~Treatment)
```

## Time series figure

The magnet effect only lasts for a couple of weeks, so I'd like to represent the fluctuations in observations as time progresses. I can put this aside the means for a more complete view of the data.

```{r time figure, echo=FALSE}
levels(brd.filt.func$Treatment)[levels(brd.filt.func$Treatment)=="B"] <- "Recent burn"
levels(brd.filt.func$Treatment)[levels(brd.filt.func$Treatment)=="C"] <- "One-year rough"

p2 <- brd.filt.func %>% 
	group_by(Treatment,Site,Week,Functional) %>% 
	summarize(Count = n()) %>% 
	group_by(Treatment, Week, Functional) %>% 
	summarize(Mean = mean(Count)) %>%
	ggplot(aes(x = Week, y = Mean, color = Functional))+
	geom_line(size=1.5, alpha=0.5)+
	geom_point(size=3.5)+
	geom_point(shape = 1,size = 3.5,colour = "black")+
	theme_bw()+
	theme(text = element_text(size = 22),
				legend.title = element_blank(),
				legend.position = 'none',
				plot.title = element_text(hjust = 0.5),
				strip.background = element_blank(),
   			strip.text.y = element_blank(),
				axis.title.x = element_text(face='bold', vjust=-2.5),
				axis.title.y = element_text(face='bold', vjust=3),
				strip.text.x = element_text(size = 18,face ='bold'),
				axis.text.x = element_text(size=10),
				legend.spacing.x = unit(0.5, 'cm'),
				plot.margin = unit(c(1,1.2,0.9,1.2),"cm"),
				legend.box.spacing = unit(1.2,'cm'),
				axis.ticks.x = element_line(size=1.2),
				axis.ticks.y = element_line(size=1.2))+
	scale_color_manual(values = c("#00AFBB", "#E7B800"))+
	scale_y_continuous(breaks=c(0,2,4,6,8,10), limits = c(0,11))+
	scale_x_date(date_breaks = "1 week", labels = function(x) format(x, "%d-%b"))+
	ylab("Mean observations")+
	xlab("Sampling period")+
	facet_wrap(~Treatment,dir = 'v')
```

## Combine the figures

```{r combined figures, warning=FALSE,message=FALSE}
library(lemon)
grid_arrange_shared_legend(p1,p2)
```

## Animated figures

Lets take a look at an animated version
```{r animated figures, warning=FALSE,message=FALSE}
library(data.table)
library(sf)
library(gganimate)
library(gifski)
library(transformr)

p3 <- brd.filt.func %>% 
	group_by(Treatment,Site,Week,Functional) %>% 
	summarize(Count = n()) %>% 
	group_by(Treatment, Week, Functional) %>% 
	summarize(Mean = mean(Count)) %>%
	ggplot(aes(x = Week, y = Mean, color = Functional))+
	geom_line(size=1.5, alpha=0.5)+
	geom_point(size=3.5)+
	transition_reveal(Week)+
	theme_bw()+
	theme(text = element_text(size = 22),
				legend.title = element_blank(),
				legend.position = 'right',
				plot.title = element_text(hjust = 0.5),
				strip.background = element_blank(),
   			strip.text.y = element_blank(),
				axis.title.x = element_text(face='bold', vjust=-2.5),
				axis.title.y = element_text(face='bold', vjust=3),
				strip.text.x = element_text(size = 18,face ='bold'),
				axis.text.x = element_text(size=15,angle = 90),
				legend.spacing.x = unit(0.5, 'cm'),
				plot.margin = unit(c(1,1.2,0.9,1.2),"cm"),
				legend.box.spacing = unit(1.2,'cm'),
				axis.ticks.x = element_line(size=1.2),
				axis.ticks.y = element_line(size=1.2))+
	scale_color_manual(values = c("#00AFBB", "#E7B800"))+
	scale_y_continuous(breaks=c(0,2,4,6,8,10), limits = c(0,11))+
	scale_x_date(date_breaks = "1 week", date_labels = "%D")+
	ylab("Mean observations")+
	xlab("Sampling period")+
	facet_wrap(~Treatment)

gganimate::animate(p3, duration = 10)
```
